Introducing Distributed Dynamic Data-intensive (D3) Science: Understanding Applications and Infrastructure

### 2015-09-12 3dpas_intro.tex
Scattered changes.

  A common feature across many science and engineering applications is the
  amount and diversity of data and computation that must be integrated to yield
  insights. Often data is increasingly large-scale and distributed. Often the
  datasets, their locations, availability, analysis and associated scheduling
  decisions are time-dependent. Collectively these give rise to dynamic
  distributed data intensive applications. This paper attempts to survey several
  such applications, to provide a common conceptual framework to understand them
  as well as examine the infrastructure used in support of applications.
  Whereas the analysis of large volumes of ``static'' data have received
  significant attention, the ability to manage large-volumes of dynamic and
  distributed data is more primitive. This work will ameliorate the state.

@@

The landscape of scientific and enterprise computing is being fundamentally
altered by the current surge in data. The resulting upheaval represents a new
reality where the volumes of data produced by scientific applications are
increasing rapidly, driven by advancing technologies (e.g., increasing compute
capacity and higher resolution sensors) and decreasing costs for data
acquisition and storage~\cite{tony-propaganda}. With the increase in data
volumes, types, and sources of data, an important concomitant development has
been an increase in the requirement for and complexity in the processing of
data. Data are often naturally distributed and have to be processed in situ, and
occasionally data that are localized have to be distributed in order to be
processed, either for reasons of performance or for collaborative analytics.
Furthermore when processing large volumes of data, data source/destination, data
production and transfer rates, and data processing capacity can all vary over
time.

The set of challenges is only partly defined by the large volumes of
data. Properties such as resource scheduling, data placement, and transfer
decisions, that were statically determinable at small scales, emerge as dynamic
and important considerations at large scales of data and distribution.

Our focus is on a subspace of the large set of problems associated with
data-intensive sciences, namely that which is characterized by the distributed
and dynamic aspects. We introduce the concept of Dynamic Distributed
Data-intensive science and refer to this subspace of data-intensive science as
D3 science.  In order to appreciate the D3 subspace we will briefly discuss the
importance and role of distribution and the pervasiveness of dynamical aspects.

@@

\section*{Acknowledgements}
This paper is an outcome of the UK e-Science Institute sponsored
Research Theme on Dynamic Distributed Data-Intensive Programming
Abstractions and Systems. This work is also sponsored by NSF
OCI-1059635.  We thank Jon Weissman and Jon Blower for useful initial
discussions, as well as 3DPAS workshop attendees. Discussions of the
selected applications benefited from the generous help of
Adam Barker,
Jon Blower,
Julian Borrill,
Silvia Delgado Olabarriaga,
Steve Fisher,
Keith Jackson,
Scott Klasky,
Bob Mann,
Don Middleton,
Kees Nieuwenhuis,
Manish Parashar,
Stephen Pascoe,
and
Jon Weissman,
though any errors are the responsibility of the authors.
Some of the work by Katz was
supported by the National Science Foundation while working at the
Foundation; any opinion, finding, and conclusions or recommendations
expressed in this material are those of the author(s) and do not
necessarily reflect the views of the National Science Foundation.

>>>>

  A common feature across many science and engineering applications is the
  amount and diversity of data and computation that must be integrated to yield
  insights. Often data is increasingly large-scale and distributed. Often the
  datasets, their locations, availability, analysis and associated scheduling
  decisions are time-dependent. Collectively these give rise to dynamic
  distributed data intensive applications.  The characteristics, requirements,
  and software systems for the analysis of large volumes of ``static'' data have
  received significant attention; dynamic, distributed and data intensive have
  received relatively less. This work will ameliorate the state by surveying
  several representative application scenarios, to provide a common conceptual
  framework to understand them as well as examine the infrastructure used in
  support of applications.

@@

The landscape of scientific and enterprise computing is being fundamentally
altered by the prodigious volumes of data.  Not only are the volumes of data
produced by scientific applications increasing rapidly, there is also a
corresponding increase in the amount of data that needs to be processed and
analyzed.  With the increase in data volumes, types, and sources of data, an
important concomitant development has been an increase in the requirement for
and complexity of the infrastructure needed to support these requirements.
Furthermore, data are often naturally distributed and have to be processed in
situ; occasionally data that are localized have to be distributed in order to be
processed, either for reasons of performance or for collaborative analytics.
The scale and significance of varying data source/destination and
transfer/production rates makes temporal variations important, which in turn
eventually impact the design and development of infrastructure to support them:
resource scheduling, data placement, and transfer decisions that were statically
determinable at small scales, emerge as dynamic and important considerations at
large scales of data and distribution.

% driven by advancing technologies (e.g., increasing compute capacity and higher
% resolution sensors) and decreasing costs for data acquisition and
% storage~\cite{tony-propaganda}.

The focus of this paper is on a subspace of the large set of problems associated
with data-intensive sciences, namely that which is characterized by {\it both}
distributed and dynamic aspects. We introduce the concept of Dynamic Distributed
Data-intensive science and refer to this subspace of data-intensive science as
D3 science.  In order to appreciate the D3 subspace we will briefly discuss the
importance and role of distribution and the pervasiveness of dynamical aspects.

@@

\section*{Acknowledgements}
This work was sponsored by NSF OCI-1059635 as well as made possible by the UK
e-Science Institute sponsored Research Theme on Dynamic Distributed
Data-Intensive Programming Abstractions and Systems. We thank Jon Weissman and
Jon Blower for useful initial discussions, as well as 3DPAS workshop
attendees. Discussions of the selected applications benefited from the generous
help of Adam Barker, Jon Blower, Julian Borrill, Silvia Delgado Olabarriaga,
Steve Fisher, Keith Jackson, Scott Klasky, Bob Mann, Don Middleton, Kees
Nieuwenhuis, Manish Parashar, Stephen Pascoe, and Jon Weissman, though any
errors are the responsibility of the authors.  Some of the work by Katz was
supported by the National Science Foundation while working at the Foundation;
any opinion, finding, and conclusions or recommendations expressed in this
material are those of the author(s) and do not necessarily reflect the views of
the National Science Foundation.
